---
title: "Biclustering and Ranklustering"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Biclustering and Ranklustering}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  out.width = "100%"
)
```

```{r library, message=FALSE, warning=FALSE}
library(exametrika)
```

## Biclustering and Ranklustering

Biclustering and Ranklustering simultaneously cluster items into fields and examinees into classes/ranks. The difference is specified via the `method` option:

- `method = "B"`: Biclustering (no filtering matrix)
- `method = "R"`: Ranklustering (with filtering matrix for ordered ranks)

### Biclustering

```{r model-biclustering, message=FALSE, warning=FALSE}
Biclustering(J35S515, nfld = 5, ncls = 6, method = "B")
```

### Ranklustering

```{r model-ranklustering, message=FALSE, warning=FALSE}
result.Ranklustering <- Biclustering(J35S515, nfld = 5, ncls = 6, method = "R")
```

```{r plot-ranklustering, fig.width=7, fig.height=5, message=FALSE, warning=FALSE}
plot(result.Ranklustering, type = "Array")
plot(result.Ranklustering, type = "FRP", nc = 2, nr = 3)
plot(result.Ranklustering, type = "RRV")
plot(result.Ranklustering, type = "RMP", students = 1:9, nc = 3, nr = 3)
plot(result.Ranklustering, type = "LRD")
```

## Finding Optimal Number of Classes and Fields

### Grid Search

`GridSearch()` systematically evaluates multiple parameter combinations and selects the best-fitting model:

```{r grid-search, message=FALSE, warning=FALSE}
result <- GridSearch(J35S515, method = "R", max_ncls = 10, max_nfld = 10, index = "BIC")
result$optimal_ncls
result$optimal_nfld
result$optimal_result
```

### Infinite Relational Model (IRM)

The IRM uses the Chinese Restaurant Process to automatically determine the optimal number of fields and classes:

```{r model-irm, message=FALSE, warning=FALSE}
result.IRM <- Biclustering_IRM(J35S515, gamma_c = 1, gamma_f = 1, verbose = TRUE)
plot(result.IRM, type = "Array")
plot(result.IRM, type = "FRP", nc = 3)
plot(result.IRM, type = "TRP")
```

## Biclustering for Polytomous Data

### Ordinal Data

```{r bic-ord, message=FALSE, warning=FALSE}
result.B.ord <- Biclustering(J35S500, ncls = 5, nfld = 5, method = "R")
result.B.ord
plot(result.B.ord, type = "Array")
```

FRP (Field Reference Profile) shows the expected score per field across latent ranks:

```{r bic-ord-frp, fig.width=7, fig.height=5, message=FALSE, warning=FALSE}
plot(result.B.ord, type = "FRP", nc = 3, nr = 2)
```

FCRP (Field Category Response Profile) shows category probabilities across ranks. The `style` parameter can be `"line"` or `"bar"`:

```{r bic-ord-fcrp, fig.width=7, fig.height=5, message=FALSE, warning=FALSE}
plot(result.B.ord, type = "FCRP", nc = 3, nr = 2)
plot(result.B.ord, type = "FCRP", style = "bar", nc = 3, nr = 2)
```

FCBR (Field Cumulative Boundary Reference) shows cumulative boundary probabilities (ordinal only):

```{r bic-ord-fcbr, fig.width=7, fig.height=5, message=FALSE, warning=FALSE}
plot(result.B.ord, type = "FCBR", nc = 3, nr = 2)
```

ScoreField and RRV plots:

```{r bic-ord-scorefield, fig.width=7, fig.height=5, message=FALSE, warning=FALSE}
plot(result.B.ord, type = "ScoreField")
plot(result.B.ord, type = "RRV")
```

### Nominal Data

```{r bic-nom, message=FALSE, warning=FALSE}
result.B.nom <- Biclustering(J20S600, ncls = 5, nfld = 4)
result.B.nom
plot(result.B.nom, type = "Array")
```

Nominal Biclustering supports FRP, FCRP, ScoreField, and RRV (but not FCBR):

```{r bic-nom-plots, fig.width=7, fig.height=5, message=FALSE, warning=FALSE}
plot(result.B.nom, type = "FRP", nc = 2, nr = 2)
plot(result.B.nom, type = "FCRP", nc = 2, nr = 2)
plot(result.B.nom, type = "FCRP", style = "bar", nc = 2, nr = 2)
plot(result.B.nom, type = "ScoreField")
plot(result.B.nom, type = "RRV")
```

## Reference

Shojima, K. (2022). *Test Data Engineering*. Springer.
