---
title: "Latent Class and Rank Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Latent Class and Rank Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  out.width = "100%"
)
```

```{r library, message=FALSE, warning=FALSE}
library(exametrika)
```

## Latent Class Analysis (LCA)

LCA classifies examinees into unordered latent classes. Specify the dataset and the number of classes.

```{r model-lca, message=FALSE, warning=FALSE}
LCA(J15S500, ncls = 5)
```

The Class Membership Matrix indicates which latent class each examinee belongs to:

```{r results-lca-membership, message=FALSE, warning=FALSE}
result.LCA <- LCA(J15S500, ncls = 5)
head(result.LCA$Students)
```

### LCA Plot Types

- **IRP**: Item Reference Profile
- **CMP**: Class Membership Profile
- **TRP**: Test Reference Profile
- **LCD**: Latent Class Distribution

```{r plot-lca, fig.width=7, fig.height=5, message=FALSE, warning=FALSE}
plot(result.LCA, type = "IRP", items = 1:6, nc = 2, nr = 3)
plot(result.LCA, type = "CMP", students = 1:9, nc = 3, nr = 3)
plot(result.LCA, type = "TRP")
plot(result.LCA, type = "LCD")
```

## Latent Rank Analysis (LRA)

LRA is similar to LCA but assumes an ordering among the latent classes (ranks). Specify the dataset and the number of ranks.

```{r model-lra, message=FALSE, warning=FALSE}
LRA(J15S500, nrank = 6)
```

Rank membership probabilities and rank-up/rank-down odds are calculated:

```{r results-lra-membership, message=FALSE, warning=FALSE}
result.LRA <- LRA(J15S500, nrank = 6)
head(result.LRA$Students)
```

```{r plot-lra, fig.width=7, fig.height=5, message=FALSE, warning=FALSE}
plot(result.LRA, type = "IRP", items = 1:6, nc = 2, nr = 3)
plot(result.LRA, type = "RMP", students = 1:9, nc = 3, nr = 3)
plot(result.LRA, type = "TRP")
plot(result.LRA, type = "LRD")
```

## LRA for Ordinal Data

LRA can also handle ordinal scale data. The `mic` option enforces monotonic increasing constraints.

```{r model-lra-ordinal, message=FALSE, warning=FALSE}
result.LRAord <- LRA(J15S3810, nrank = 3, mic = TRUE)
```

Score-rank relationship visualizations:

```{r plot-lra-ordinal-score, fig.width=7, fig.height=5, message=FALSE, warning=FALSE}
plot(result.LRAord, type = "ScoreFreq")
plot(result.LRAord, type = "ScoreRank")
```

Item-rank relationship plots:

- **ICBR**: Item Category Boundary Reference -- cumulative probability curves for each category threshold
- **ICRP**: Item Category Response Profile -- probability of each response category across ranks

```{r plot-lra-ordinal-icbr, fig.width=7, fig.height=5, message=FALSE, warning=FALSE}
plot(result.LRAord, type = "ICBR", items = 1:4, nc = 2, nr = 2)
plot(result.LRAord, type = "ICRP", items = 1:4, nc = 2, nr = 2)
```

Rank membership profiles for individual examinees:

```{r plot-lra-ordinal-rmp, fig.width=7, fig.height=5, message=FALSE, warning=FALSE}
plot(result.LRAord, type = "RMP", students = 1:9, nc = 3, nr = 3)
```

## LRA for Rated/Nominal Data

For multiple-choice tests (nominal scale), LRA can analyze response patterns including distractor choices.

```{r lra-rated, message=FALSE, warning=FALSE}
result.LRArated <- LRA(J35S5000, nrank = 10, mic = TRUE)
```

```{r plot-lra-rated-score, fig.width=7, fig.height=5, message=FALSE, warning=FALSE}
plot(result.LRArated, type = "ScoreFreq")
plot(result.LRArated, type = "ScoreRank")
```

```{r plot-lra-rated-icrp, fig.width=7, fig.height=5, message=FALSE, warning=FALSE}
plot(result.LRArated, type = "ICRP", items = 1:4, nc = 2, nr = 2)
```

## Reference

Shojima, K. (2022). *Test Data Engineering*. Springer.
